Contention Management for 5G

### 2017-03-14 article/article.tex

    \begin{proposition}
    There is an algorithm, which finds a $(P,\tau)$-periodic affectation in time $O(n^2)$ when $P \geq 3n\tau$.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive macro slots of time of size $\tau$. The algorithm works in the following way: select the first route $k+1$ for which the offset as not been chosen. Try every of the $3N-k$ offsets which put the message in one free macro slot in the forward period. It also fixes the position of a message in the backward period, and we choose the first offset which does not create a collision in the backward period. 
     Assume we choose the offset of the route $r_{k+1}$, we have  at least $3n - k$ free macro-slots, since $P \geq 3n\tau$. Each of these $3n - k$ possible offset values translates into $3n - k$ positions of messages in the backward period. All these positions are separated by at least $\tau$ slots. There are already $k$ messages of size $\tau$ in the backward period. One such message can intersect at most $2k$ potential positions, therefore  amongst the possible $3n - k$ positions, there are  $3n - k -2k$ which are without collision. Since $k < n$, one of the choice is without collision, which proves that the algorithm terminates and find a
     $(P,\tau)$-periodic affectation. 
     \end{proof}
     This algorithm works in a complexity $n^2$, since we have to try every free macro-slots for every routes.
      \begin{center}
      \includegraphics[scale=0.3]{ex3nt.png}
      \end{center}
    \todo{Avec une structure de données plus maligne on doit pouvoir faire n, non ?}
	\begin{algorithm}[H]
	\caption{Greedy affectation}
	\begin{algorithmic}
	\REQUIRE ${\cal R}_{\cal C}$, period $P$
	\ENSURE A P-periodic affectation in p $\leq P$, or FAILURE
	\STATE $P1[P/\tau]$ slots of size $\tau$ in forward period.
	\STATE $P2[P]$ slots backward period.
	\FORALL{source $s_i$ in S }

@@

	\IF{$P1[j]$ is free AND the corresponding area in P2 is free}

@@

	\IF{No intervals are found for i}
	\STATE return FAILURE
	\ENDIF
	\ENDFOR

@@

    \subsubsection{Exhaustive generation}
	\begin{algorithm}[H]
	\caption{Exhaustive Generation}  
	\begin{algorithmic}
	\REQUIRE A routage graph ${\cal R}_{\cal C}$, period $P$, packet size $\tau$
	\ENSURE $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$
	\STATE Forward-budget $\leftarrow$ $P$ - n * $\tau$
	\STATE Backward-budget $\leftarrow$ $P$ - n * $\tau$
	\STATE Free-Intervals $\leftarrow$ list of free intervals, init to $[0;P[$
	\FORALL{route i in  ${\cal R}_{\cal C}$}
	\FORALL{j in Free-Intervals }
	\IF{Message of the route i does not collides with scheduled routes}
	\STATE $m_{s_i} \leftarrow $ the first slot of Free-Intervals[j]
	\STATE Split the Free-Intervals considering the new packet
	\STATE Forward-budget $\leftarrow$ Forward-budget - {\em lost size}
	\STATE Backward-budget $\leftarrow$ Backward-budget - {\em lost size}
	\STATE call Exhaustive Generation on remaining routes
	\ENDIF
	\ENDFOR
	\ENDFOR

      \end{algorithmic}
      \end{algorithm}
      {\bf {\em Lost size}}, is the size lost when an interval is too few to put another packet. Ex : if the interval [250;3250[ receives a packet of size 2500, there is 500 slots lost, because there is no way to uses the slots for another packet.\\

      This algorithm enumerates all the solutions by traversing a tree. The leaves of the tree are 
      the $(P\tau)$-periodic affectations, and the nodes are partial solutions. A partial solution is a choice of a starting time for a subset of the routes.
      For each route, the algorithm tries every possible starting offset until it finds one available (that allows the message to come back without collisions). When a route have been scheduled, it take another route that have not been scheduled yet.
      This creates a new sub-tree in which the algorithm will try every possible starting offset too, and create a sub-tree on the following route too.
      If no offset are available for a route, this means that the partial solution is not good, so the algorithm backtracks over the tree that is,
      it goes back to the father of the current sub-tree.
      Once a leaf is obtained, the algorithm stops and return the scheduling.\\      
      To reduce the number of useless calculation, in each time we add another packet, we consider the lost size in both periods, thus, we know when a period has no many great enough free area to put a packet.

>>>>

    \begin{proposition}
    There is an algorithm, which finds a $(P,\tau)$-periodic affectation in time $O(n^2)$ when $P \geq 3n\tau$.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive macro slots of time of size $\tau$. The algorithm works in the following way: select the first route $k+1$ for which the offset as not been chosen. Try every of the $3N-k$ offsets which put the message in one free macro slot in the forward period. It also fixes the position of a message in the backward period, and we choose the first offset which does not create a collision in the backward period. 
     Assume we choose the offset of the route $r_{k+1}$, we have  at least $3n - k$ free macro-slots, since $P \geq 3n\tau$. Each of these $3n - k$ possible offset values translates into $3n - k$ positions of messages in the backward period. All these positions are separated by at least $\tau$ slots. There are already $k$ messages of size $\tau$ in the backward period. One such message can intersect at most $2k$ potential positions, therefore  amongst the possible $3n - k$ positions, there are  $3n - k -2k$ which are without collision. Since $k < n$, one of the choice is without collision, which proves that the algorithm terminates and find a
     $(P,\tau)$-periodic affectation. 
     \end{proof}
     This algorithm works in a complexity $n^2$, since we have to try every free macro-slots for every routes.
      \begin{center}
      \includegraphics[scale=0.3]{ex3nt.png}
      \end{center}
    \todo{Avec une structure de données plus maligne on doit pouvoir faire n, non ?}
	\begin{algorithm}[H]
	\caption{Greedy affectation}
	\begin{algorithmic}
	\REQUIRE ${\cal R}_{\cal C}$, period $P$
	\ENSURE A P-periodic affectation in p $\leq P$, or FAILURE
	\STATE $P1[P/\tau]$ macro slots of size $\tau$ in forward period.
	\STATE $P2[P]$ slots backward period.
	\FORALL{source $s_i$ in S }

@@

	\IF{$P1[j]$ is free AND the corresponding slots in P2 are free}

@@

	\IF{No intervals are found for $s_i$}
	\STATE return FAILURE
	\ENDIF
	\ENDFOR

@@

    \subsubsection{Exhaustive generation}
	\begin{algorithm}[H]
	\caption{Exhaustive Generation}  
	\begin{algorithmic}
	\REQUIRE A routage graph ${\cal R}_{\cal C}$, period $P$, packet size $\tau$
	\ENSURE $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$
	\STATE Forward-budget $\leftarrow$ $P$ - n * $\tau$
	\STATE Backward-budget $\leftarrow$ $P$ - n * $\tau$
	\STATE Free-Intervals $\leftarrow$ list of free intervals in the backward period, init to $[0;P[$
	\FORALL{source $s_i$ in S}
	\FORALL{j in Free-Intervals }
	\IF{Message of the route $r_{s_i}$ does not collides with scheduled routes}
	\STATE $m_{s_i} \leftarrow $ the first slot of Free-Intervals[j]
	\STATE Split the Free-Intervals considering the new packet
	\STATE Forward-budget $\leftarrow$ Forward-budget - {\em lost size}
	\STATE Backward-budget $\leftarrow$ Backward-budget - {\em lost size}
	\STATE call Exhaustive Generation on remaining routes
	\ENDIF
	\ENDFOR
	\ENDFOR

      \end{algorithmic}
      \end{algorithm}
      {\bf {\em Lost size}}, is the size lost when an interval is too few to put another packet. Ex : if the interval [250;3250[ receives a packet of size 2500, there is 500 slots lost, because there is no way to uses the slots for another packet.\\
      \todo{ecrire l'algo? si ou le réécrire}

      This algorithm enumerates all the solutions by traversing a tree. The leaves of the tree are 
      the $(P\tau)$-periodic affectations, and the nodes are partial solutions. A partial solution is a choice of a starting time for a subset of the routes.
      For each route, the algorithm tries every possible starting offset until it finds one available (that allows the message to come back without collisions). When a route have been scheduled, it take another route that have not been scheduled yet.
      This creates a new sub-tree in which the algorithm will try every possible starting offset too, and create a sub-tree on the following route too.
      If no offset are available for a route, this means that the partial solution is not good, so the algorithm backtracks over the tree that is,  it goes back to the father of the current sub-tree, and continue from there.
      Once a leaf is obtained, the algorithm stops and return the $(P\tau)$-periodic affectation.\\      
      To reduce the number of useless calculation, we can imagine cuts in the tree :
      \begin{enumerate}
       \item In the forward windows, each message is placed after the previous one, that is $m_{s_i} > m_{s_j}$ (here, $s_i$ is the source taken first in the algorithm, the order is not fixed). Thus, if the number of the remaining slots in the forward period is inferior than the number of slots needed to put the remaining messages, we can consider that the actual partial solution is not good.
       \item  Each time we add a new message, we cut a free interval in two new free intervals, plus the scheduled message. Each free intervals has a number of macro slots (ex: 2,6), and if the number of all macro slots is inferior than the number of remaining messages, the actual partial solution is not good.
       \item ...
      \end{enumerate}
